\clearpage

\chapter{Annotated Bibliography}

\cite{zirui} Li, Z., Faheem, F., and Husung, S. ''Collaborative Model-Based Systems Engineering Using Dataspaces and SysML v2'' Systems (2024)

\begin{itemize}
    \item This paper explores how to enable collaborative, multi-company MBSE by combining SysML v2 modeling with a secure framework for data exchange. The proposed solution relies on a modeling guideline to maintain consistency and interoperability, which is why its relevant as it justifies why standardized guidelines are essential for leveraging SysML v2's new features in a collaborative, multi-tool environment.
\end{itemize}

\cite{bergemann} Bergemann, S. ''Challenges in Multi-View Model Consistency Management for Systems Engineering.'' Modellierung 2022 Satellite Events (2022).

\begin{itemize}
    \item This paper investigates the critical challenge of managing model consistency across multiple, heterogeneous viewpoints in MBSE. It analyzes existing approaches against a set of industrial requirements, identifying significant gaps in current methods. This work is essential as it defines the problem of inconsistency that new, formalized SysML v2 guidelines are intended to solve, thereby motivating Hypotheses 1 and 2.
\end{itemize}

\cite{boelsen} Boelsen, K., et al. ''SysML v2 based modelling guidelines for mechanical system elements.'' Forschung im Ingenieurwesen (2025)

\begin{itemize}
    \item Test
\end{itemize}

\cite{eduardo} Cibrián, E., et al. ''Ensuring Semantic Consistency in SysML v2 Models Through Metamodel-Driven Validation.'' IEEE Access (2025)

\begin{itemize}
    \item Test
\end{itemize}

\cite{dehn} Dehn, S., et al. ''Enhancing model-based development with formalized requirements: integrating temporal logic and SysML v2 for comprehensive state and transition modeling'' Forschung im Ingenieurwesen (2025).

\begin{itemize}
    \item Test
\end{itemize}

\cite{yildirim} Yildirim, U., et al. ''Flow Heuristics for Functional Modelling in Model-Based Systems Engineering.'' Proc. Int. Conf. on Engineering Design (ICED23) (2023)

\begin{itemize}
    \item Test
\end{itemize}

\cite{friedenthal} Friedenthal, S. ''Future Directions for MBSE with SysML v2.'' Proc. 11th Int. Conf. on Model-Based Software and Systems Engineering (2023)

\begin{itemize}
    \item Test
\end{itemize}

\cite{jansen} Jansen, N., et al. ''The Language of SysML v2 under the Magnifying Glass.'' Journal of Object Technology (2022)

\begin{itemize}
    \item Test
\end{itemize}

\cite{beers} Beers, L., et al. ''A SysML Profile for the Standardized Description of Processes during System Development.'' arXiv (2024)

\begin{itemize}
    \item Test
\end{itemize}

\addtocontents{toc}{\vspace{0.8cm}}